### 2018-08-14 ICRA19/covControl.tex
New ICRA paper

\subsection{Using boundaries: stable configurations of a swarm}\label{subsec:FluidInTank}
One method to control a swarm's shape in a bounded workspace is to simply push in a given direction until the swarm conforms to the boundary. Like fluid settling in a tank, the stable final configuration minimizes potential energy.
\paragraph{Square workspace}
%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: 
We first examine the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a large swarm of robots as they move inside a square workspace under the influence of a force pulling in the direction $\beta$. The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$, $A\in [0,1]$. Under a global input, the swarm moves to a side of the workspace and forms a polygonal shape that minimizes potential energy, as shown in Fig.~\ref{fig:SquareFill},  see also \cite{Zhao2016mathematicaSquare}.

@@

Computing means, variances, covariance, and correlation requires integrating over $R$, the region containing the swarm:  %This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else. 
%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: %, integrated over the unit square with $x$ and $y$ from 0 to 1:

@@

The region of integration $R$ is the polygon containing the swarm. For example, if $A<1/2$ and the force angle is $\beta$, the mean when $R$ is a triangular region in the lower-left corner is:
\begin{align}\label{eq:meanInSquareWorkspaceLL}
\bar{x}(A,\beta) &= \frac{\int_0^{\sqrt{2 A \tan (\beta)}} \left(\int_0^{\cot (\beta) \left(\sqrt{2 A \tan (\beta)}-x\right)} \, dy\right) x \, dx}{A} \nonumber \\
	&=\frac{ \sqrt{2}}{3} \sqrt{A \tan (\beta )}\\
\bar{y}(A,\beta) &= \frac{\int_0^{\sqrt{2 A \tan (\beta)}} \left(\int_0^{\cot (\beta) \left(\sqrt{2 A \tan (\beta)}-x\right)} y \, dy\right) \, dx}{A} \nonumber\\
	&=\frac{\sqrt{2}}{3}  \sqrt{A \cot (\beta )}.
\end{align}
The full equations are included in the appendix~\cite{2016arXiv160901830S}, and are summarized in Fig.~\ref{fig:SquareFill}. A few highlights are that the correlation is maximized $\pm$1/2 when the swarm is in a triangular shape. The covariance of a triangle is always $\pm(A/18)$. Variance is minimized in the direction of $\beta$ and maximized orthogonal to $\beta$ when the swarm is in a rectangular shape. The range of mean positions are maximized when $A$ is small.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Circular workspace}
Though rectangular boundaries are common in artificial workspaces, biological workspaces are usually rounded.
Similar calculations can be made for a circular workspace.  The workspace is a circle centered at (0,0) with radius 1 and thus area $\pi$.
For notational simplicity, the swarm is parameterized by the global control input signal $\beta$ and the fill-level $h$.  
Under a global input, the robot swarm fills the region under a chord with area
\begin{align}
A(h) = \arccos(1-h)-(1-h) \sqrt{(2-h) h}.
\end{align}
For a circular workspace, the locus of mean positions are aligned with $\beta$ and the mean position is at radius $r(h)$ from the center:
\begin{align}
r(h) = \frac{2 (-(h-2) h)^{3/2}}{3 \left(\sqrt{-(h-2) h} (h-1)+\arccos(1-h)\right)}.
\end{align}
Variance $\sigma^2_x(\beta,h)$ is maximized at $\beta = \pi/2+n \pi$ and $h\approx1.43$, while covariance is maximized at $\beta = \pi3/4+n \pi$ and $h\approx0.92.$ For small $h$ values, correlation approaches $\pm1$. Results are displayed in Fig.~\ref{fig:CircleFill}, see also \cite{Zhao2016mathematica}.

@@

Let the control input be a vector force $\vec{F}$ with magnitude $F$ and orientation $\theta$ with respect to a line perpendicular to and into the nearest boundary. $N$ is the normal or perpendicular force between the robot and the boundary. The force of friction $F_f$ is nonzero if the robot is in contact with the boundary and  $\sin(\theta) < 0$. The resulting net force on the robot, $F_{\text{\emph{forward}}}$, is aligned with the wall and given by
\begin{align}
F_{\text{\emph{forward}}} &=  F \sin(\theta) - F_f  \nonumber \\
\text{where }  F_f &= \begin{cases}  \mu_f N, &  \mu_f N < F \sin(\theta)  \label{eq:frictionmodel}  \\
F \sin(\theta), & \text{else} \end{cases} \\ %\sign(F \sin(\theta) ) \cdot  \max(0, | F sin\theta |- |F_f|)
\text{and } N &= F \cos(\theta) .\nonumber
\end{align}
 Fig.~\ref{fig:friction} shows the resultant forces on two robots when one is touching a wall. 
Though each receives the same inputs,  they experience different net forces.
  For ease of analysis, the following algorithms assume $\mu_f$ is infinite and robots touching the wall are prevented from sliding along the wall.
This means that if one robot is touching the wall and another robot is free, the touching robot will not move when the control input is parallel or into the wall. 
There are many alternate models of friction that also break control symmetry. Fig.~\ref{fig:friction}c shows fluid flow along a boundary.  Fluid in the free-flow region moves uniformly, but flow decreases to zero in the boundary layer \cite{fluidMechanics}.  Force in such a system can be calculated as
\begin{align}
%u(y) = u_0 \left(1- \frac{(y-h)^2}{h^2}\right) = u_0 \frac{y}{h} \left(2- \frac{y}{h}\right) \label{eq:boundarylayerflow} THiS EQUATION IS TERRIBLE! PLOT IT.
F_{forward}(y) &= F - F_f\begin{cases}  \frac{h-y}{h}   , &  y<h \label{eq:boundarylayerflow} \\
0, & \text{else} \end{cases}.
\end{align}

@@

\subsection{Maximizing correlation using wall friction}\label{subsec:ClosedLoopCovarianceControl}
Using the environment to individually move $n$ particles to goal positions in Alg. \ref{alg:PosControlNRobots} requires $O(n^2)$ time, while  Alg. \ref{alg:optimalAlg} can only control two particles. The controllers in Section \ref{sec:theory} are efficient because they require only a single move but the range of possible configurations are limited. This section presents an efficient technique to generate desired correlations.

Assume an obstacle-free, bounded, unit-size, square workspace. 
 As shown in Fig.~\ref{fig:SquareFill}, the maximum correlation occurs when the swarm is pushed in the direction $\beta = 3\pi/4$. 
 This correlation as a function of swarm area $A$ is never larger than 1/2, and the maximum correlation decays to 0 as $A$ grows to 1. By  \eqref{eq:covAndcorrInSquareWorkspace}, this maximum correlation is:
\begin{align} \label{eq:GravityCorrelation}
\rho_{xy} =  \begin{cases}  \frac{1}{2}  , &  0\le A\le \frac{1}{2}  \\
% \frac{3 A (2 (A-2) A+1)}{4 A \left(A^2+3 \sqrt{2-2 A}-6\right)-12 \sqrt{2-2 A}+17}-1
 \frac{3 A (2 (A-2) A+1)}{4 A^3-24 A+\sqrt{2} (12 A-12) \sqrt{1-A}+17}-1
 , & \frac{1}{2} \le A\le 1
\end{cases}.
\end{align}
 If friction obeys the linear boundary layer model of \eqref{eq:boundarylayerflow} with boundary layer thickness $h$ and maximum friction $F_f$ equal to the maximum applied force $F$, we can generate  larger correlations.
 If the swarm size is smaller than $\approx 0.43$ and the boundary layer is sufficiently thick we can generate correlations larger than 1/2  using boundary friction.

@@

\subsection{Efficient control of correlation}
%A set of simulations were conducted to demonstrate the importance of boundary friction.  
This section examines maximum correlation values as a function of  $w,L$ using \eqref{eq:correlationFriction}
from Section  \ref{subsec:ClosedLoopCovarianceControl}. 
The maximum correlation using boundary layer friction $\displaystyle  \max_{w,L} \left( \rho(A,h,w,L) \right)$ can be found by gradient descent, as shown in Fig.~\ref{fig:ControlledCorrelation}. 
For swarms with small area, this method enables generating the full range of correlations $\pm 1$, while the stable configuration method from Section \ref{sec:theory} could only generate correlations $\pm 1/2$.  As  the swarm area $A$ increases above $\approx 0.43$, the stable configuration method is more effective.
Larger boundary layers $h$ enable more control of correlation.
The multimedia attachment shows efficient control of correlation with simulations using the 2D physics engine Box2D~\cite{catto2010box2d} and 144 disc-shaped robots with boundary layer model \eqref{eq:boundarylayerflow}.

>>>>

\subsection{Using boundaries: stable configurations of a swarm}\label{subsec:FluidInTank}
One method to control a swarm's shape in a bounded workspace is to simply push in a given direction until the swarm conforms to the boundary. Like fluid settling in a tank, the stable final configuration minimizes potential energy.
\paragraph{Square workspace}
%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: 
We first examine the mean $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ of a large swarm of robots as they move inside a square workspace under the influence of a force pulling in the direction $\beta$. The swarm is large, but the robots are small in comparison, and together occupy a constant area $A$, $A\in [0,1]$. Under a global input, the swarm moves to a side of the workspace and forms a polygonal shape that minimizes potential energy, as shown in Fig.~\ref{fig:SquareFill}. % see also \cite{Zhao2016mathematicaSquare}.

@@

Computing means, variances, covariance, and correlation requires integrating over $R$, the region containing the swarm and are calculated as%This is simplified using an indicator function $\bm{1}_A(x,y)$ that returns 1 if inside the region containing the swarm, and 0 else. 
%The formulas for means $(\bar{x},\bar{y})$, covariance $(\sigma^2_x,\sigma^2_y,\sigma_{xy})$, and correlation $\rho_{xy}$ are as follows: %, integrated over the unit square with $x$ and $y$ from 0 to 1:

@@

The region of integration $R$ is the polygon containing the swarm. For example, if $A<1/2$ and the force angle is $\beta$, the mean when $R$ is a triangular region in the lower-left corner is:
\begin{align}\label{eq:meanInSquareWorkspaceLL}
\bar{x}(A,\beta) &= \frac{\int_0^{\sqrt{2 A \tan (\beta)}} \left(\int_0^{\cot (\beta) \left(\sqrt{2 A \tan (\beta)}-x\right)} \, dy\right) x \, dx}{A} \nonumber \\
	&=\frac{ \sqrt{2}}{3} \sqrt{A \tan (\beta )},\\
\bar{y}(A,\beta) &= \frac{\int_0^{\sqrt{2 A \tan (\beta)}} \left(\int_0^{\cot (\beta) \left(\sqrt{2 A \tan (\beta)}-x\right)} y \, dy\right) \, dx}{A} \nonumber\\
	&=\frac{\sqrt{2}}{3}  \sqrt{A \cot (\beta )}.
\end{align}
The full equations are summarized in Fig.~\ref{fig:SquareFill}. A few highlights are that the correlation is maximized $\pm$1/2 when the swarm is in a triangular shape. The covariance of a triangle is always $\pm(A/18)$. Variance is minimized in the direction of $\beta$ and maximized orthogonal to $\beta$ when the swarm is in a rectangular shape. The range of mean positions are maximized when $A$ is small.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Circular workspace}
Though rectangular boundaries are common in artificial workspaces, biological workspaces are usually rounded.
Similar calculations can be made for a circular workspace.  The workspace is a circle centered at (0,0) with radius 1 and thus area $\pi$.
For notational simplicity, the swarm is parameterized by the global control input signal $\beta$ and the fill-level $h$.  
Under a global input, the robot swarm fills the region under a chord with area
\begin{align}
A(h) = \arccos(1-h)-(1-h) \sqrt{(2-h) h}.
\end{align}
For a circular workspace, the locus of mean positions are aligned with $\beta$ and the mean position is at radius $r(h)$ from the center which is
\begin{align}
r(h) = \frac{2 (-(h-2) h)^{3/2}}{3 \left(\sqrt{-(h-2) h} (h-1)+\arccos(1-h)\right)}.
\end{align}
Variance $\sigma^2_x(\beta,h)$ is maximized at $\beta = \pi/2+n \pi$ and $h\approx1.43$, while covariance is maximized at $\beta = \pi3/4+n \pi$ and $h\approx0.92.$ For small $h$ values, correlation approaches $\pm1$. Results are displayed in Fig.~\ref{fig:CircleFill}.%, see also \cite{Zhao2016mathematica}.

@@

Let the control input be a vector force $\vec{F}$ with magnitude $F$ and orientation $\theta$ with respect to a line perpendicular to and into the nearest boundary. $N$ is the normal or perpendicular force between the robot and the boundary. The force of friction $F_f$ is nonzero if the robot is in contact with the boundary and  $\sin(\theta) < 0$. The resulting net force on the robot, $F_{\text{\emph{forward}}}$, is aligned with the wall and given by
\begin{align}
F_{\text{\emph{forward}}} &=  F \sin(\theta) - F_f , \nonumber \\
\text{where }  F_f &= \begin{cases}  \mu_f N, &  \mu_f N < F \sin(\theta)  \label{eq:frictionmodel}  \\
F \sin(\theta), & \text{else} \end{cases} \\ %\sign(F \sin(\theta) ) \cdot  \max(0, | F sin\theta |- |F_f|)
\text{and } N &= F \cos(\theta) .\nonumber
\end{align}
 Fig.~\ref{fig:friction} shows the resultant forces on two robots when one is touching a wall. 
Though each receives the same inputs,  they experience different net forces.
  For ease of analysis, the following algorithms assume $\mu_f$ is infinite and robots touching the wall are prevented from sliding along the wall.
This means that if one robot is touching the wall and another robot is free, the touching robot will not move when the control input is parallel or into the wall. 
There are many alternate models of friction that also break control symmetry. Fig.~\ref{fig:friction}c shows fluid flow along a boundary.  Fluid in the free-flow region moves uniformly, but flow decreases to zero in the boundary layer \cite{fluidMechanics}.  Force in such a system can be calculated as
\begin{align}
%u(y) = u_0 \left(1- \frac{(y-h)^2}{h^2}\right) = u_0 \frac{y}{h} \left(2- \frac{y}{h}\right) \label{eq:boundarylayerflow} THiS EQUATION IS TERRIBLE! PLOT IT.
F_{forward}(y) &= F - F_f\begin{cases}  \frac{h-y}{h}   , &  y<h \label{eq:boundarylayerflow} \\
0, & \text{else} \end{cases}.
\end{align}

@@

\subsection{Maximizing correlation using wall friction}\label{subsec:ClosedLoopCovarianceControl}
%Using the environment to individually move $n$ particles to goal positions in Alg. \ref{alg:PosControlNRobots} requires $O(n^2)$ time, while  Alg. \ref{alg:optimalAlg} can only control two particles. The controllers in Section \ref{sec:theory} are efficient because they require only a single move but the range of possible configurations are limited. This section presents an efficient technique to generate desired correlations.

Assume an obstacle-free, bounded, unit-size, square workspace. 
 As shown in Fig.~\ref{fig:SquareFill}, the maximum correlation occurs when the swarm is pushed in the direction $\beta = 3\pi/4$. 
 This correlation as a function of swarm area $A$ is never larger than 1/2, and the maximum correlation decays to 0 as $A$ grows to 1. By  \eqref{eq:covAndcorrInSquareWorkspace}, this maximum correlation is
\begin{align} \label{eq:GravityCorrelation}
\rho_{xy} =  \begin{cases}  \frac{1}{2}  , &  0\le A\le \frac{1}{2}  \\
% \frac{3 A (2 (A-2) A+1)}{4 A \left(A^2+3 \sqrt{2-2 A}-6\right)-12 \sqrt{2-2 A}+17}-1
 \frac{3 A (2 (A-2) A+1)}{4 A^3-24 A+\sqrt{2} (12 A-12) \sqrt{1-A}+17}-1
 , & \frac{1}{2} \le A\le 1
\end{cases}.
\end{align}
 If friction obeys the linear boundary layer model of \eqref{eq:boundarylayerflow} with boundary layer thickness $h$ and maximum friction $F_f$ equal to the maximum applied force $F$, we can generate  larger correlations.
 If the swarm size is smaller than $\approx 0.43$ and the boundary layer is sufficiently thick we can generate correlations larger than 1/2  using boundary friction.

@@

\subsection{Efficient control of correlation}
%A set of simulations were conducted to demonstrate the importance of boundary friction.  
This section examines maximum correlation values as a function of  $w,L$ using \eqref{eq:correlationFriction}
from Section  \ref{subsec:ClosedLoopCovarianceControl}. 
The maximum correlation using boundary layer friction $\displaystyle  \max_{w,L} \left( \rho(A,h,w,L) \right)$ can be found by gradient descent, as shown in Fig.~\ref{fig:ControlledCorrelation}. 
For swarms with small area, this method enables generating the full range of correlations $\pm 1$.%, while the stable configuration method from Section \ref{sec:theory} could only generate correlations $\pm 1/2$.  
 As  the swarm area $A$ increases above $\approx 0.43$, the stable configuration method is more effective.
Larger boundary layers $h$ enable more control of correlation.
The multimedia attachment shows efficient control of correlation with simulations using the 2D physics engine Box2D~\cite{catto2010box2d} and 144 disc-shaped robots with boundary layer model \eqref{eq:boundarylayerflow}.

### 2018-08-14 ICRA19/intro.tex
New ICRA paper

 Positioning is a foundational capability for a robotic system, e.g. placement of brachytherapy seeds. 
 In previous work, we showed that the 2D position of each particle in such a swarm is controllable if the workspace contains a single obstacle the size of one particle \cite{AaronManipulation2013}.
 However, requiring a single, small, rigid obstacle suspended in the middle of the workspace is often an unreasonable constraint, especially in 3D.
This paper relaxes that constraint, and provides position control algorithms that only require non-slip wall contacts.
We assume that particles in contact with the boundaries have zero velocity if the uniform control input pushes the particle into the wall.

\begin{figure}
\centering
\vspace{1.5em}
%\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
\begin{overpic}[width=0.45\columnwidth]{firstpicLeft.pdf}\put(28,-10){workspace}\end{overpic}
\begin{overpic}[width=0.45\columnwidth]{magneticsetup.pdf}\put(22,-8){magnetic setup}\end{overpic}
\vspace{1em}
\caption{\label{fig:IntroPic}
Workspace and magnetic setup for an experiment to position particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
} \vspace{-1em}
\end{figure}
%\todo{add the picture of magnetic setup}

The paper is arranged as follows. 
After a review of recent related work in Sec.  \ref{sec:RelatedWork},
 %Sec. \ref{subsec:FluidInTank} provides analytical position control results of stable configurations in two canonical workspaces with frictionless walls.  These results are limited in the set of shapes that can be generated.  To extend the range of possible shapes,
  Sec.  \ref{sec:theory} introduces a  model for boundary interaction and two shortest path results for representative workspaces.   
We provide an algorithm to arbitrarily position two particles in Sec.  \ref{sec:PostionControl2Robots}.
Section  \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.  \ref{sec:expResults} describes hardware experiments, as shown in Fig.~\ref{fig:IntroPic}. 
 We end with directions for future research in Sec.  \ref{sec:conclusion}.

This paper is an elaboration of preliminary work in a conference paper \cite{shahrokhi2017algorithms} which considered only square workspaces. This work extends the analysis to convex workspaces and 3D positioning. This paper also implements the algorithms using a hardware setup inspired by the anatomy of the gastrointestinal tract.

>>>>

 Positioning is a foundational capability for a robotic system, e.g. placement of brachytherapy seeds. 
 In previous work, we provided position control algorithms that only require non-slip wall contacts to position two robots \cite{shahrokhi2017}
We assumed that particles in contact with the boundaries have zero velocity if the uniform control input pushes the particle into the wall. This work analyses interactions between the boundary and the particle swarm to control shape of the swarm. In \cite{shahrokhi2018TRO} we showed how to control swarm statistics including mean position and variance. This work extends that work to covariance control.

%    \begin{figure}
%    \centering
%    \vspace{1.5em}
%    %\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
%    \begin{overpic}[width=0.45\columnwidth]{firstpicLeft.pdf}\put(28,-10){workspace}\end{overpic}
%    \begin{overpic}[width=0.45\columnwidth]{magneticsetup.pdf}\put(22,-8){magnetic setup}\end{overpic}
%    \vspace{1em}
%    \caption{\label{fig:IntroPic}
%    Workspace and magnetic setup for an experiment to position particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
%    } \vspace{-1em}
%    \end{figure}
%\todo{add the picture of magnetic setup}

The paper is arranged as follows. 
After a review of recent related work in Sec.  \ref{sec:RelatedWork},
 %Sec. \ref{subsec:FluidInTank} provides analytical position control results of stable configurations in two canonical workspaces with frictionless walls.  These results are limited in the set of shapes that can be generated.  To extend the range of possible shapes,
  Sec.  \ref{sec:angle} introduces angle of repose, a parameter of the particle swarm that can be used for shape control. 
Section  \ref{sec:covControl} describes implementations of the algorithms of covariance control in simulation and hardware experiments. 
 % Sec.  \ref{sec:expResults} 
 We end with directions for future research in Sec.  \ref{sec:conclusion}.

### 2018-08-14 ICRA19/relatedWork.tex
New ICRA paper

If a control input causes the particles to collide with obstacles at different times, inverting the control input does not undo the action. 
 Due to this lack of time-reversibility, techniques that require a bidirectional graph, e.g. PRM \cite{kavraki1996probabilistic} and RRT* \cite{lavalle2006planning} are not suitable.
  Instead, this paper employs a greedy search strategy.

>>>>

%If a control input causes the particles to collide with obstacles at different times, inverting the control input does not undo the action. 
% Due to this lack of time-reversibility, techniques that require a bidirectional graph, e.g. PRM \cite{kavraki1996probabilistic} and RRT* \cite{lavalle2006planning} are not suitable.
%  Instead, this paper employs a greedy search strategy.

### 2018-08-14 ICRA19/theory.tex
New ICRA paper

Consider a swarm of granular particles applying force to a rod. 
If the rod moves slower than the particles, these particles will build up behind the rod in a characteristic triangular shape defined by an apex angle. This piling up is common to all granular media, and the angle formed is a function of the \emph{angle of repose}. Three different values of angle of repose is shown in Fig.~\ref{fig:angle}. The center of mass of the rod is in the middle of the rod, but center of mass of the granular particles changes for different values of angle of repose. % for the media, see \cite{angleofRepose}. 
 By measuring the angle of repose for the particles shown in the top plot of Fig.~\ref{fig:AngleOfReposeForce}, we can estimate the force and torque that the swarm is applying to the rod as a function of the rod's length, the angle of repose, and orientation of the rod.
 We define the angle of repose as $\alpha$, the rod's orientation relative to 90$^\circ$ from the particle movement vector as $\theta$, and the rod's length as $\ell$. 
 \begin{figure}
\centering
\renewcommand{\figwid}{\columnwidth}
\begin{overpic}[width =\figwid]{Angle.pdf}%\put(1,55){a)}
\end{overpic}
\caption{\label{fig:angle} Different values of angle of repose is shown when granular particles move faster than the rod.
}
\end{figure}
By integrating over the triangular shape, the force applied to the rod (when a unit area of particles produces 1 N of force) is:
\begin{figure}
\centering
\renewcommand{\figwid}{\columnwidth}
\begin{overpic}[width =0.6\figwid]{AngleOfRepose.pdf}%\put(1,55){a)}
\end{overpic}\\
\vspace{0.5em}
\begin{overpic}[width =0.6\figwid]{AngleOfReposeForce.pdf}%\put(1,55){a)}
\end{overpic}\\
\vspace{0.5em}
\begin{overpic}[width =0.6\figwid]{AngleOfReposeTorque.pdf}%\put(1,55){a)}
\end{overpic}
\vspace{-0.5em}
\caption{\label{fig:AngleOfReposeForce} Top plot shows colored particulate heaped up on pink-colored long rods. 
%Particulate is moving in the $-y$ direction, and the rods are tilted at $\theta=\{-45,-22.5,0,22.5,45\}^\circ$ with respect to the $x$ axis. 
% A thin gray line extends upwards from the rod COM, showing how more particulate is heaped on the right side of the rod for positive $\theta$ and more on the left side for negative $\theta$. 
%  This uneven particulate generates a restoring torque. 
 Middle plot shows the force applied to the rod and bottom the torque as a function of $\theta$ for four angle of repose values.
 %The rod length $\ell=1$. 
   The maximum torque values are shown with black dots.
% Generating code is in the attachment. 
%\vspace{-2em}
}
\end{figure}

@@

The force for different angle of repose values are shown in the middle plot of Fig.~\ref{fig:AngleOfReposeForce}. 
Torque will also be similarly defined as
\begin{align}
\tau(\theta, \alpha, \ell) =\left\{
\begin{array}{ll}
\frac{\ell^3\Big(\cos(2\alpha)-\cos(2\theta)\Big)\sin{\theta} }{48\sin^2\alpha}&   -\alpha<\theta<\alpha\\
0 &    \textrm{otherwise .}\\
\end{array} 
\right.
\end{align}
Torque is shown in the bottom plot of Fig.~\ref{fig:AngleOfReposeForce}.
 Given sufficient particles to pile up to the angle of repose, this torque tends to stabilize the object to be perpendicular to the pushing direction.
Force is maximized with $\theta=0$, but the $\theta$ value that maximizes torque is a function of $\alpha$:
\begin{align}
\theta_{t_{\max}} = \frac{\sin(\alpha)}{\sqrt{3}}.
\end{align}
To maximize the torque a particulate swarm applies on a thin rod, the swarm should move in the direction $-\theta_{t_{\max}} - 90^\circ$ with respect to the long axis of the rod.

>>>>

Consider a swarm of granular particles applying force to a rod. 
If the rod moves slower than the particles, these particles will build up behind the rod in a characteristic triangular shape defined by an apex angle. This piling up is common to all granular media, and the angle formed is a function of the \emph{angle of repose}. Three different values of angle of repose is shown in Fig.~\ref{fig:angle}. The center of mass of the rod is in the middle of the rod, but center of mass of the granular particles changes for different values of angle of repose. % for the media, see \cite{angleofRepose}. 
 By measuring the angle of repose for the particles shown in the top plot of Fig.~\ref{fig:AngleOfReposeForce}, we can estimate the force and torque that the swarm is applying to the rod as a function of the rod's length, the angle of repose, and orientation of the rod.
 We define the angle of repose as $\alpha$, the rod's orientation relative to 90$^\circ$ from the particle movement vector as $\theta$, and the rod's length as $\ell$. 
 \begin{figure}
\centering
\renewcommand{\figwid}{\columnwidth}
\begin{overpic}[width =\figwid]{Angle.pdf}%\put(1,55){a)}
\end{overpic}
\caption{\label{fig:angle} Different values of angle of repose is shown when granular particles move faster than the rod.
}
\end{figure}
By integrating over the triangular shape, the force applied to the rod (when a unit area of particles produces 1 N of force) is
\begin{figure}
\centering
\renewcommand{\figwid}{\columnwidth}
\begin{overpic}[width =0.6\figwid]{AngleOfRepose.pdf}%\put(1,55){a)}
\end{overpic}\\
\vspace{0.5em}
\begin{overpic}[width =0.6\figwid]{AngleOfReposeForce.pdf}%\put(1,55){a)}
\end{overpic}\\
\vspace{0.5em}
\begin{overpic}[width =0.6\figwid]{AngleOfReposeTorque.pdf}%\put(1,55){a)}
\end{overpic}
\vspace{-0.5em}
\caption{\label{fig:AngleOfReposeForce} Top plot shows colored particulate heaped up on pink-colored long rods. 
%Particulate is moving in the $-y$ direction, and the rods are tilted at $\theta=\{-45,-22.5,0,22.5,45\}^\circ$ with respect to the $x$ axis. 
% A thin gray line extends upwards from the rod COM, showing how more particulate is heaped on the right side of the rod for positive $\theta$ and more on the left side for negative $\theta$. 
%  This uneven particulate generates a restoring torque. 
 Middle plot shows the force applied to the rod and bottom the torque as a function of $\theta$ for four angle of repose values.
 %The rod length $\ell=1$. 
   The maximum torque values are shown with black dots.
% Generating code is in the attachment. 
%\vspace{-2em}
}
\end{figure}

@@

The force for different angle of repose values are shown in the middle plot of Fig.~\ref{fig:AngleOfReposeForce}. 
Torque will also be similarly defined as
\begin{align}
\tau(\theta, \alpha, \ell) =\left\{
\begin{array}{ll}
\frac{\ell^3\Big(\cos(2\alpha)-\cos(2\theta)\Big)\sin{\theta} }{48\sin^2\alpha}&   -\alpha<\theta<\alpha\\
0 &    \textrm{otherwise .}\\
\end{array} 
\right.
\end{align}
Torque is shown in the bottom plot of Fig.~\ref{fig:AngleOfReposeForce}.
 Given sufficient particles to pile up to the angle of repose, this torque tends to stabilize the object to be perpendicular to the pushing direction.
Force is maximized with $\theta=0$, but the $\theta$ value that maximizes torque is a function of $\alpha$ and is defined as
\begin{align}
\theta_{t_{\max}} = \frac{\sin(\alpha)}{\sqrt{3}}.
\end{align}
To maximize the torque a particulate swarm applies on a thin rod, the swarm should move in the direction $-\theta_{t_{\max}} - 90^\circ$ with respect to the long axis of the rod.
